---
title: "Final Project"
author: "Christian Tillich"
date: "October 21, 2017"
output: pdf_document
---

```{r startup, include=FALSE, warning=FALSE}
library(rjags)
library(jagstools)
library(knitr)
library(mnlogit)
library(ordinal)
knitr::opts_chunk$set(warning = FALSE, results="asis", cache = TRUE)

build.model <- function(
   model
  ,data
  ,inits
  ,params
  ,n.iter = 25000
  ,...
){
  M <- jags.model(textConnection(paste(model, collapse="\n")), data=data,inits=inits, quiet=T,...)
  R <- coda.samples(M,params,n.iter=n.iter)
  R
}
```

```{r data}
df <- read.table("../data/tb.txt",header=T) 
```

```{r jags_model}
code <- "
  model{
    for(i in 1:134){
      #Core Model
      reactors[i] ~ dpois(mu[i])
      mu[i] <- adar[i]*nu[i]
      log(nu[i]) <- beta.0 + beta.sex*sex[i] + beta.age[age[i]] + beta.type[type[i]]

      #Deviance Calc
      dev[i] <- 2*(reactors[i]*log(reactors[i]/mu[i]) - (reactors[i] - mu[i]))
    } 

    tot.dev <- sum(dev[])

    #Priors
    beta.0 ~ dnorm(0, 0.001)
    beta.sex ~ dnorm(0, 0.001)
    for(j in 2:3){beta.age[j] ~ dnorm(0, 0.001)}
    beta.age[1] <- 0
    for(j in 2:4){beta.type[j] ~ dnorm(0, 0.001)}
    beta.type[1] <- 0
  }
" %>% strsplit('\n') %>% unlist

inits <- list(
   beta.0 = 0
  ,beta.sex = 0
  ,beta.age = c(NA, 0, 0)
  ,beta.type = c(NA, 0, 0, 0)
)

params <- c('beta.0','beta.sex','beta.age[2:3]','beta.type[2:4]','tot.dev')
monitor <- c('beta.0','beta.sex','beta.age','beta.type','tot.dev')
build.model(code, df, inits, params, n.chains=2, n.adapt=500, n.iter=25000) %>%
  jagsresults(monitor) %>%
  kable
```


# Question 1

Total Deviance is~355. So (a) above 300. 

# Question 2

`beta.type[2]` has a confidence interval of roughly [0.63, 1.56]. This interval
is all positive, suggesting sufficient evidence that Cervids bear increased risk. 


# Binomial Refactor

```{r q3}
code %>% 
  {.[5] <- "      reactors[i] ~ dnegbin(p[i], theta)"; .} %>%
  append('      p[i] <- theta/(theta + mu[i])',5) %>%
  append('    theta ~ dunif(0,10)',22) %>%
  build.model(df, inits, c(params, 'theta'), n.chains=2, n.adapt=500, n.iter=25000) %>%
  jagsresults(c(monitor,'theta')) %>%
  kable
```

# Question 3

After rephrasing as a negative-binomial model, none of the type variables 
have sufficient evidence to be considered significant. 

# Question 4

The posterior mean for theta is ~0.56

