---
title: "Homework 1"
author: "Christian Tillich"
date: "September 24, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
library(rjags)
library(knitr)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, results="asis")
```



# Setup 

```{r p1_setup, results='hide'}

code <- list(
   'data {n0 <- 0.1}'
  ,'model{ '
  ,'  for (i in 1:10) {y[i] ~ dnorm(mu,tau)} '
  ,'  tau ~ dgamma(1,0.01)'
  ,'  sigma2 <- 1/tau' 
  ,'  sigma <- sqrt(sigma2)'
  ,'  mu ~ dnorm(0, n0*tau)'
  ,'}'
  #,sep = '\n'
)

build.model <- function(
   jags.program
  ,data = list(y=c(10,9,9,8,9.5,7,12,11,8,10.5))
  ,inits = list(list(mu=10,tau=1),list(mu=9, tau=0.1))
  ,params = c("mu","sigma","sigma2")
  ,n.iter=25000
  ,...
){

  mdl <- textConnection(paste(jags.program, collapse='\n'))
  
  M <- jags.model(mdl,data=data, inits = inits, quiet = TRUE, ...) #n.chains=2,n.adapt=500)
  R <- coda.samples(M,params,n.iter=n.iter)
  
  return(R)
}

show.model <- function(jags.model){
  summary(jags.model)$statistics %>% as.data.frame() %>% kable %>% print
  summary(jags.model)$quantiles %>% as.data.frame() %>% kable %>% print
}
```

# Problem 1

```{r p1}
build.model(code, n.chains=2, n.adapt = 500) %>% 
  show.model
```

The lower point of the 95% confidence interval is 1.125, so below 2. 


# Problem 2

```{r p2}
#Specifying the JAGS code here. 
append(code,'pt.sigma.gt <- step(sigma - 2.5)',6) %>%
  build.model(n.chains=2, n.adapt = 500, params = c('mu','sigma','pt.sigma.gt')) %>%
  show.model
```

The posterior probability of sigma > 2.5 is 0.035, so above 0.02. 

# Problem 3

```{r p3}
#Specifying the JAGS code here. 
append(code,'cv <- sigma / mu',6) %>%
  build.model(n.chains=2, n.adapt = 500, params = c('mu','sigma','cv')) %>%
  show.model
```

The 95% confidence interval for the coefficient of variation is [0.120, 0.285], 
so the upper point is above 0.25. 

# Problem 4

```{r p4}
#Specifying the JAGS code here. 
append(code,'cv <- sigma / mu',6) %>%
  append('p.cv.gt <- step(cv - 0.2)', 7) %>% 
  build.model(n.chains=2, n.adapt = 500, params = c('mu','sigma','cv', 'p.cv.gt')) %>%
  show.model
```

The probability of the cv > 0.20 is 27%. So that probability is below 30%. 

# Problem 5

```{r p5}
new.data <- list(y=c(1,9,9,8,9.5,7,12,11,8,10.5))

add.code <- list(
   'for (i in 1:10) {'
  ,'y[i] ~ dnorm(mu, tau)'
  ,'yrep[i] ~ dnorm(mu, tau)'
  ,'p.rep.gt[i] <- step(yrep[i] - y[i])'
  ,'}'
)

append(code[-3], add.code, 3) %>%
  build.model(
     data = new.data
    ,n.chains=2
    ,n.adapt = 500
    ,params = c('mu','sigma','p.rep.gt')
  ) %>%
  show.model
```

1) Observation 1 has the highest probability of y_rep > y, at p = 98.7%. 
2) That probability is over 95%. 
3) It's a problematic model. Per the model, the probability of recreating the 
first observation is < 2%. 


# Problem 6

```{r p6}
append(code[-3], add.code, 3) %>%
  {.[5] <- 'y[i] ~ dt(mu, tau, 4)'; .} %>%
  build.model(
     data = new.data
    ,n.chains = 2
    ,n.adapt = 500
    ,params = c('mu','sigma','p.rep.gt')
  ) %>% 
  show.model
```

1) Observeration 1 still has the highest probability of y_rep > y, now at p = 
99.8%
2) It's still over 95%
3) The mean went from 8.41 in Problem 5 to 8.91 in Problem 6. But that would
help explain why the it became less likely to replicate observation 1 despite 
having the fatter tails (not gonna lie, this still seems counter-intuitive)